What we {\sf eval} in the shadows

### 2021-04-01 paper/main.tex
minor

\mypara{Intercession.} A common use case for \eval is to be combined
with \c{match.call}. \c{match.call} walks up the call stack, captures
the code that invoked the currently executing function, and returns it
as an unevaluated expression. The pattern is to transform a call to
some function \c{f} into a call to \c{g} with some arguments retained
and others modified. As an illustration, consider the \c{vcpart}
package's function \c{tvcglm} that is translated to a call to \c{tvcm}
with two modifications to the argument list: argument \c{control}
can't be missing and \c{fit} is set  \c{"glm"}. The
function ends with a call to \c{eval.parent} to ensure that the
rewritten call is evaluted in the same environment the original call
was.
\begin{lstlisting}
 tvcglm <- function(formula, data, family, control=tvcglm_control(), ...) {
   Call <- match.call()
   Call[[1L]] <- as.name("tvcm")
   if (!"control" %in% names(Call)) Call$control<-formals(tvcglm)$control
   if ("fit" %in% names(list(...))) warning("'fit' is ignored. ")
   Call$fit <- "glm"
   return(eval.parent(Call))
 }
\end{lstlisting}
This pattern is recognizable by the fact that the expression is a call
and the target environment is that of the parent.

\mypara{Code Generation.} The more traditional use of \eval is to
execute code that was assembled by the programmer into a string. Here
we show the method \c{plot} for class \c{sback} in package
\c{wsbackfit}, simplified for explanatory purposes. The function takes
a long argument list, the names of which are captured in the list
\c{opt}. The string \c{stub} is composed of a subset of the arguments
passed to this function; the variable is used to construct a call to
\c{base::plot} which will draw a plot. \NOTE{This example is a bit
  messed up. What is var? Does the use of ... in the eval grab the
  arguments of this call? If yes does it not end up with xlab twice?
  HELP} The \parse and \eval combination is used by the the \source
function in the base library to load R code from a file in the current
workspace.
\begin{lstlisting}
 plot.sback <- function(x,...) {
   opt <- names(list(...))
   stub <- paste(
    ifelse("xlab" %in% opt,"",paste(",xlab=\"",var,"\"",sep="")),
    ifelse("main" %in% opt,"",main.aux),
    ifelse("type" %in% opt,"",",type=\"l\""),
    sep = "")
   plot <- paste("plot(x.data,",stub,",...)",sep="")
   eval(parse(text=plot))
   ...
\end{lstlisting}
This pattern is recognizable by its use of \c{parse} to turn a string
into an expression.

\mypara{Debloating.} \Eval is often used as a means to reduce
boilerplate code; simple and repetitive code can easily be replaced
with judcious use of \eval. For example, the \c{data.table} package
uses \eval to calls the \c{options} function with named arguments
taken from a vector of strings. While the benefits are limited in this
example, it is an attractive tool for programmers.
\begin{lstlisting}
  opts = c("datatable.verbose"="FALSE", # ... many others
  for (i in names(opts))
    eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}
This pattern is a special case of code generation, recognizable by the
fact that \eval is executed in a loop.

>>>>

\mypara{Intercession.} A common use case for \eval is to be combined with
\c{match.call}. \c{match.call} walks up the call stack, captures the code that
invoked the currently executing function, and returns it as an unevaluated
expression. The pattern is to transform a call to some function \c{f} into a
call to \c{g} with some arguments retained and others modified. As an
illustration, consider the \c{vcpart} package's function \c{tvcglm} that is
translated to a call to \c{tvcm} with two modifications to the argument list:
argument \c{control} can't be missing and \c{fit} is set \c{"glm"}. The function
ends with a call to \c{eval.parent} to ensure that the rewritten call is
evaluted in the same environment the original call was.
\begin{lstlisting}
 tvcglm <- function(formula, data, family, ...) {
   k <- match.call()
   k[[1]] <- as.name("tvcm")
   if ("fit" %in% names(list(...))) warning("'fit' ignored.")
   k$fit <- "glm"
   eval.parent(k)
 }
\end{lstlisting}
This pattern is recognizable by the fact that the expression is a call
and the target environment is that of the parent.

\mypara{Code Generation.} The more traditional use of \eval is to execute code
that was assembled by the programmer into a string. Here we show the method
\c{plot} for class \c{sback} in package \c{wsbackfit}, simplified for
explanatory purposes. The function takes a long argument list, the names of
which are captured in the list \c{opt}. The string \c{stub} is composed of a
subset of the arguments passed to this function; the variable is used to
construct a call to \c{base::plot} which will draw a plot. \NOTE{This example is
  a bit messed up. What is var? Does the use of ... in the eval grab the
  arguments of this call? If yes does it not end up with xlab twice? HELP} The
\parse and \eval combination is used by the the \source function in the base
library to load R code from a file in the current workspace.
\begin{lstlisting}
 plot.sback <- function(x,...) {
   opt <- names(list(...))
   stub <- paste(
    ifelse("xlab" %in% opt,"",paste(",xlab=\"",var,"\"",sep="")),
    ifelse("type" %in% opt,"",",type=\"l\""), sep="")
   plot <- paste("plot(x.data,",stub,",...)",sep="")
   eval(parse(text=plot))
\end{lstlisting}
This pattern is recognizable by its use of \c{parse} to turn a string
into an expression.

\mypara{Debloating.} \Eval is often used as a means to reduce
boilerplate code; simple and repetitive code can easily be replaced
with judcious use of \eval. For example, the \c{data.table} package
uses \eval to calls the \c{options} function with named arguments
taken from a vector of strings. While the benefits are limited in this
example, it is an attractive tool for programmers.
\begin{lstlisting}
  opts = c("datatable.verbose"="FALSE", # ...many others
  for (i in names(opts))
    eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}
This pattern is a special case of code generation, recognizable by the
fact that \eval is executed in a loop.
